### 2016-08-25 rebuttal.tex
Start working on rebuttal

\documentclass[12pt]{article}

\usepackage{pscyr}

\usepackage[T2A]{fontenc}
\usepackage[cp1251]{inputenc}

\usepackage[english,russian]{babel}

\begin{document}

\begin{center}
  \Large Ответ рецензенту
\end{center}

\end{document}